---
title: "Sentiment Analysis"
author: "Will Schrepferman"
date: "3/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(rvest)
library(gt)
library(janitor)
library(ggplot2)
library(tidytext)
library(glue)
library(stringr)
library(plotly)
```

## Milestone 5

For ease of use, I decided to use the same github repo I used last time and modify the plot that I created!

This file is largely the same as ms_5, except with a lot more detail and care put into the actual plot.

```{r basefile, include = FALSE}
## THIS IS PARTIALLY RECREATED FROM A KAGGLE TUTORIAL, LINK IS ABOVE

# get a list of input files

files <- list.files("input")

readfile <- function(file){
  
    # get the file
  
    fileName <- glue("input/", file, sep = "")
    
    # get rid of any trailing spaces
    
    fileName <- trimws(fileName)

    # read in the new file
    
    fileText <- glue(read_file(fileName))
    
    # remove any dollar signs (they're special characters in R)
    
    fileText <- gsub("\\$", "", fileText)
    
    # create the tibble using the file's text, file name, year by grabbing the last four digits of file name, and president
    
    base <- tibble(text = fileText, file = file, year = as.numeric(str_match(file, "\\d{4}")), president = str_match(file, "(.*?)_")[2])

    # return that tibble
    
    return(base)
}

# create a blank tibble

base_file <- tibble()

# run through all files in input

for(i in files){

    # found out about Rbind on stackoverflow, it was applicable here
  
    base_file <- rbind(base_file, readfile(i))
}

```

```{r webdata, include = FALSE}

# Read in some more state of the union data from the web

# function to get the rightmost values in a character string

substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}


# website with information on word counts of all state of the union addresses

web_data <- read_html("https://www.presidency.ucsb.edu/documents/presidential-documents-archive-guidebook/annual-messages-congress-the-state-the-union-3")

# scrape table using xpath

raw_data <- web_data %>%
  html_nodes(xpath = "/html/body/div[2]/div[4]/div/section/div/section/div/div/div/div[2]/table") %>%
  html_table()

# ensure it is a df

final_data <- raw_data[[1]]

# remove the header row

final_data <- final_data[2:nrow(final_data), ]

# clean up data by renaming headers and filtering out columns that are averages

final_data_clean <- final_data %>%
  rename("president" = "X1", "date" = "X2", "delivery" = "X3", "words" = "X4") %>%
  select(date, delivery, words) %>%
  filter(delivery != "average") %>%
  filter(delivery != "average spoken") %>%
  filter(delivery != "average written") %>%
  filter(delivery != "no annual messages") %>%
  filter(delivery != "average (ex. 1981)") %>%
  filter(delivery != "average (ex. 1989)") %>%
  filter(delivery != "average (ex. 1993)") %>%
  filter(delivery != "average (ex. 2001)") %>%
  filter(delivery != "average (ex. 2009)") %>%
  filter(delivery != "average (ex. 2017)") %>%
  
  # add a year column from the date column by using the substring function at the beginning of this code chunk
  
  mutate(year = as.numeric((substrRight(date, 4))))
```

```{r combinedata, include = FALSE}

# combine web data with input data

comb_data <- base_file %>%
  inner_join(final_data_clean, by = "year")
```

```{r sentiments, include = FALSE}

# function designed to get the sentiment of any file in our SOTU inputs

GetSentiment <- function(file){
  
    # beginning of this file is exact same as earlier readfile function
  
    fileName <- glue("input/", file, sep = "")
    fileName <- trimws(fileName)
    fileText <- glue(read_file(fileName))
    fileText <- gsub("\\$", "", fileText)

    # break down text into individual words
    
    tokens <- tibble(text = fileText) %>% unnest_tokens(word, text)
    
    # get the total number of words
    
    num_words <- tokens %>%
      count()

    # get the sentiment from the text:
    
    sentiment <- tokens %>%
      
      # uses the existing Bing sentiment lexicon, more on this in the Kaggle link
      
      inner_join(get_sentiments("bing")) %>%
      
      # count the number of positive & negative words
      
      count(sentiment) %>%
      
      # make data wide rather than narrow
      
      spread(sentiment, n, fill = 0) %>%
      
      # number of positive words minus number of negative words divided by total words
      
      mutate(sentiment = ((positive - negative)/as.double(num_words))) %>%
      
      # add file name, year, and president like I did before
      
      mutate(file = file) %>%
      mutate(year = as.numeric(str_match(file, "\\d{4}"))) %>%
      mutate(president = str_match(file, "(.*?)_")[2])

    # return our sentiment tibble
    
    return(sentiment)
}

# exactly as before, make a blank tibble and read in results with rbind

sentiments <- tibble()

for(i in files){
    sentiments <- rbind(sentiments, GetSentiment(i))
}


# disambiguate Bush Sr. and George W. Bush
# correct president in applicable rows

bushSr <- sentiments %>%
  filter(president == "Bush") %>%
  filter(year < 2000) %>%
  mutate(president = "Bush Sr.")

# remove incorrect rows

sentiments <- anti_join(sentiments, sentiments[sentiments$president == "Bush" & sentiments$year < 2000, ])

# add corrected rows to data_frame

sentiments <- full_join(sentiments, bushSr)

# same process but for John Adams and John Quincy Adams

quincyAdams <- sentiments %>%
  filter(president == "Adams") %>%
  filter(year > 1824) %>%
  mutate(president = "Quincy Adams")

sentiments <- anti_join(sentiments, sentiments[sentiments$president == "Adams" & sentiments$year > 1824, ])

sentiments <- full_join(sentiments, quincyAdams)

# same process but for FDR and Teddy Roosevelt

FDR <- sentiments %>%
  filter(president == "Roosevelt") %>%
  filter(year > 1932) %>%
  mutate(president = "F. Roosevelt")

sentiments <- anti_join(sentiments, sentiments[sentiments$president == "Roosevelt" & sentiments$year > 1932, ])

sentiments <- full_join(sentiments, FDR)

sentiments <- sentiments %>%
  mutate(president = ifelse(president == "Roosevelt", "T. Roosevelt", president))

```

```{r sentiment_plot, include = FALSE}
## THIS IS WHERE MY ACTUAL NEW CODE IS


# filter for only speeches from the last 100 years (expanded slightly to include all of Wilson's term)

sentiments <- sentiments %>%
  filter(year >= 1913)

# manually add in the party of each president
# I could have found an online table, but that would honestly
# have taken longer than manually inputting this minute amount of data

sentiments <- sentiments %>%
  mutate(Party = "party") %>%
  mutate(Party = ifelse(president == "Wilson", "Democrat", Party)) %>%
  mutate(Party = ifelse(president == "Harding", "Republican", Party)) %>%
  mutate(Party = ifelse(president == "Coolidge", "Republican", Party)) %>%
  mutate(Party = ifelse(president == "Hoover", "Republican", Party)) %>%
  mutate(Party = ifelse(president == "F. Roosevelt", "Democrat", Party)) %>%
  mutate(Party = ifelse(president == "Truman", "Democrat", Party)) %>%
  mutate(Party = ifelse(president == "Eisenhower", "Republican", Party)) %>%
  mutate(Party = ifelse(president == "Kennedy", "Democrat", Party)) %>%
  mutate(Party = ifelse(president == "Johnson", "Democrat", Party)) %>%
  mutate(Party = ifelse(president == "Nixon", "Republican", Party)) %>%
  mutate(Party = ifelse(president == "Ford", "Republican", Party)) %>%
  mutate(Party = ifelse(president == "Carter", "Democrat", Party)) %>%
  mutate(Party = ifelse(president == "Reagan", "Republican", Party)) %>%
  mutate(Party = ifelse(president == "Bush Sr.", "Republican", Party)) %>%
  mutate(Party = ifelse(president == "Clinton", "Democrat", Party)) %>%
  mutate(Party = ifelse(president == "Bush", "Republican", Party)) %>%
  mutate(Party = ifelse(president == "Obama", "Democrat", Party)) %>%
  mutate(Party = ifelse(president == "Trump", "Republican", Party)) %>%
  
  # turn year into numeric
  
  mutate(year = as.numeric(year))

# make plot in the same way as last time

sentiment_plot <- ggplot(sentiments, aes(x = year, y = sentiment)) +
  labs(x = "Year", y = "Positive Sentiment", title = "Positive Sentiment of Modern State of the Union Addresses") +
  geom_point(aes(color = Party, president = president)) +
  
  # add colors by party, not by president
  scale_color_manual(values = c("Democrat" = "blue", "Republican" = "red")) +
  
  # manually input vertical lines to disambiguate between presidents
  
  geom_vline(xintercept = 1920.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1922.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1928.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1932.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1945.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1953.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1961.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1963.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1969.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1974.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1977.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1981.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1988.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 1992.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 2000.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 2008.5, linetype="dotted", color = "darkgrey", size= .4) +
  geom_vline(xintercept = 2016.5, linetype="dotted", color = "darkgrey", size= .4)

# allows user to hover over points to see which president it is

interactive <- ggplotly(sentiment_plot, tooltip = c("president", "year", "sentiment"))

```

## The Graph

Here's the completed graph!

Note on methodology: In order to avoid trickiness with political parties outside of the modern two-party system, I used only speeches from the last 100 years, expanded slightly to include the entirety of Wilson's term. "Positive sentiment" refers to the value of positive tone detected in the speech using sentiment analysis, with 0 being a totally neutral speech.

Note on graph use: It's interactive! Hover over each dot to see which president it is. The vertical lines mark different presidents!

```{r}
interactive
```


Fun historical notes that my analysis picks up on: it appears that Eisenhower was certifiably bubbly, with consistently more positive speeches than other presidents. FDR delivered the only two overall tonally negative speeches in 1938 and 1942- not surprising that the more negative speeches came during bleak times like the tail end of the Great Depression and the beginning of World War 2.

Also, another interesting trend worth potentially examining more: eyeballing it, it looks like most presidents get more positive with time. Step 1: become President. Step 2: ?. Step 3: happiness.


